---
title: "ggplot2 extensions"
execute:
  echo: true
  cache: false
  message: false
  warning: false
format: 
  revealjs:
    width: 1152
    height: 648
    highlight-style: joyce.theme
    theme: [simple, custom3.scss]
    scrollable: true
    chalkboard: true
    slide-number: true
    code-fold: true

knitr:
  opts_chunk: 
    fig.width: 6
    fig.height: 4
    fig.align: center
    out.width: 100%
    
language:
  code-summary: ""
---    


# Classifying by construct:

GGPLOT2 GRAMMAR

## New Geoms

`ggforce` -- arcs, bezier curves, splines, voronai, hulls, delaunay, and others

`ggalt` -- splines, encircling points, dumbbell and lollipop charts, and others

`ggsignif` -- significance brackets with p values

`geomtextpath` -- curved text

## `ggforce`

```{r}
#| fig-width: 12
#| fig-height: 4
library(ggforce)
library(patchwork)
df <- data.frame(r = 1:48, 
                 start = runif(48, 0, 2*pi), 
                 end = runif(48, 0, 2*pi))

p1 <- ggplot(df, aes(x0 = 0, y0 = 0, r = r, 
               start = start, end = end,
               color = r)) + 
  geom_arc() + 
#  coord_fixed() + 
  scale_color_viridis_c() +
  labs(title = "geom_arc()") +
  theme_bw() +
  theme(legend.position = "none",
        title = element_text(family = "courier", face = "bold"))

control_points <- data.frame(
  x = c(0, 2, 3, 5, 1, 1.25, 1.25, 3.75, 3.5, 4.25, 4.5),
  y = c(4, 5, 3, 4.25, 1, 1.5, 4.25, 3.75, 2, 0, 1.5),
  group = c(1, 1, 1, 1, 2, 2, 2, 3, 3, 3, 3)
)

p2 <- ggplot(control_points, aes(x = x, y = y, group = group, color = factor(group))) +
  geom_bezier(linewidth = 2, alpha = .75) +
  geom_point() +
#  coord_fixed() +
  scale_x_continuous(name = NULL) +
  scale_y_continuous(name = NULL, breaks = 0:5) +
  labs(title = "geom_bezier()") +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        title = element_text(family = "courier", face = "bold"))

p3 <- ggplot(iris, aes(Petal.Length, Petal.Width, color = Species)) + 
  geom_point() + 
  geom_mark_ellipse(aes(label = Species), label.fontsize = 8) +
  scale_x_continuous(expand = c(.2, .2)) +
  scale_y_continuous(expand = c(.2, .2)) +
  labs(title = "geom_mark_ellipse()", x = NULL, y = NULL) +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        plot.title = element_text(family = "courier", face = "bold"))

p1 + p2 + p3
```

## `ggalt`

```{r}
#| fig-width: 12
#| fig-height: 4
library(ggplot2)
library(ggalt)

dat <- data.frame(x=c(1:10, 1:10, 1:10),
                  y=c(sample(15:30, 10), 2*sample(15:30, 10), 3*sample(15:30, 10)),
                  group=factor(c(rep(1, 10), rep(2, 10), rep(3, 10)))
)


p1 <- ggplot(dat, aes(x, y, group=group, color=factor(group))) +
    geom_point(color="black") +
    geom_smooth(se=FALSE, linetype="dashed", size=0.5) +
    geom_xspline(spline_shape=-0.4, size=0.5) +
    labs(title = "geom_xspline()") +
    theme_bw() +
    theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        title = element_text(family = "courier", face = "bold"))

d <- data.frame(x=c(1,1,2),y=c(1,2,2)*100)

p2 <- ggplot(d,aes(x,y))
p2 <- p2 + scale_x_continuous(expand=c(0.5,1))
p2 <- p2 + scale_y_continuous(expand=c(0.5,1))

p2 <- p2 + geom_encircle(s_shape=0.5, expand=0.1, colour="red") + geom_point() +
  labs(title = "geom_encircle()") +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        title = element_text(family = "courier", face = "bold"))

df <- data.frame(trt=LETTERS[1:5], l=c(20, 40, 10, 30, 50), r=c(70, 50, 30, 60, 80))

p3 <- ggplot(df, aes(y=trt, x=l, xend=r)) + 
  geom_dumbbell(size=3, color="#e3e2e1", 
                colour_x = "#5b8124", colour_xend = "#bad744",
                dot_guide=TRUE, dot_guide_size=0.25) +
  labs(x=NULL, y=NULL, title="geom_dumbbell()") +
  theme(panel.grid.major.x=element_line(size=0.05)) +
  theme(panel.grid.major.y=element_blank()) +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        title = element_text(family = "courier", face = "bold"))

p1 + p2 + p3
```

:::{.pt5}
Code source: [https://github.com/hrbrmstr/ggalt](https://github.com/hrbrmstr/ggalt)
:::

## `ggsignif`

```{r}
#| fig-width: 12
#| fig-height: 4
library(ggplot2)
library(ggsignif)

p1 <- ggplot(mpg, aes(class, hwy)) +
  geom_boxplot() +
  geom_signif(
    comparisons = list(c("compact", "midsize"), c("minivan", "suv")),
    map_signif_level = TRUE, textsize = 6
  ) +
  ylim(NA, 48) +
  labs(title = "geom_signif()") +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        title = element_text(family = "courier", face = "bold"))

p2 <- ggplot(diamonds, aes(x = cut, y = carat)) +
  geom_boxplot(aes(fill = color)) +
  geom_signif(comparisons = list(
    c("Fair", "Good"),
    c("Very Good", "Ideal")
  ), textsize = 3) +
  facet_wrap(~color) +
  ylim(NA, 6.3) +
  labs(title = "compatible with facets") +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        title = element_text(family = "courier", face = "bold")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p1 + p2
```

:::{.pt5}
Code source: [https://github.com/const-ae/ggsignif](https://github.com/const-ae/ggsignif)
:::

## `geomtextpath`

```{r}
#| fig-width: 12
#| fig-height: 4
library(geomtextpath)

t <- seq(5, -1, length.out = 1000) * pi

spiral <- data.frame(x = sin(t) * 1:1000, 
                     y = cos(t) * 1:1000,
                     text = paste("Like a circle in a spiral,",
                                  "like a wheel within a wheel,",
                                  "never ending or beginning,",
                                  "on an ever spinning reel")
                     )

p1 <- ggplot(spiral, aes(x, y, label = text)) +
  geom_textpath(size = 4, vjust = 2, text_only = TRUE) +
  coord_equal(xlim = c(-1500, 1500), ylim = c(-1500, 1500)) +
    labs(title = "geom_textpath()") +
    theme_bw() +
    theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        title = element_text(family = "courier", face = "bold"))

df <- data.frame(x = spline(1:5, runif(5), xout = seq(1, 5, 1/100))$y,
                 y = spline(1:5, runif(5), xout = seq(1, 5, 1/100))$y,
                 z = "A curved textbox on an arbitrary path")

p2 <- ggplot(df, aes(x, y, label = z)) + geom_labelpath(size = 4, fill = "#F6F6FF") +
    labs(title = "geom_labelpath()") +
    theme_bw() +
    theme(legend.position = "none",
      panel.grid.minor = element_blank(),
      title = element_text(family = "courier", face = "bold"))

p3 <- ggplot(iris, aes(x = Sepal.Length, colour = Species, label = Species)) +
  geom_textdensity(size = 6, fontface = 2, hjust = 0.2, vjust = 0.3) +
  theme(legend.position = "none") +
  labs(title = "geom_textdensity()") +
  theme_bw() +
  theme(legend.position = "none",
      panel.grid.minor = element_blank(),
      title = element_text(family = "courier", face = "bold"))

p1 + p2 + p3
```

:::{.pt5}
Code source: [https://allancameron.github.io/geomtextpath/](https://allancameron.github.io/geomtextpath/)
:::

## New Scales

`ggsci` -- color palettes inspired by colors used in scientific journals, data visualization libraries, science fiction movies, and TV shows

`ggnewscale` -- use multiple fill/color scales in one plot

`ggbreak` -- set breakpoints for truncating the plot

## `ggsci`

```{r}
#| fig-width: 12
#| fig-height: 4
library(ggsci)

data("diamonds")

p1 <- ggplot(
  subset(diamonds, carat >= 2.2),
  aes(x = table, y = price, colour = cut)
) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "loess", alpha = 0.05, linewidth = 1, span = 1)

p1_npg <- p1 + scale_color_npg() +
  labs(title = "scale_color_npg()") +
  theme_bw() +
  theme(legend.position = "none",
      panel.grid.minor = element_blank(),
      title = element_text(family = "courier", face = "bold"))

p1_rickandmorty <- p1 + scale_color_rickandmorty() +
  labs(title = "scale_color_rickandmorty()") +
  theme_bw() +
  theme(legend.position = "none",
      panel.grid.minor = element_blank(),
      title = element_text(family = "courier", face = "bold"))

data("mtcars")
cor <- cor(unname(mtcars))
cor_melt <- data.frame(
  Var1 = rep(seq_len(nrow(cor)), times = ncol(cor)),
  Var2 = rep(seq_len(ncol(cor)), each = nrow(cor)),
  value = as.vector(cor)
)

p3 <- ggplot(cor_melt, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(colour = "black", linewidth = 0.3) +
  theme_void() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )

p3_gsea <- p3 + scale_fill_gsea() +
  labs(title = "scale_fill_gsea()") +
  theme_bw() +
  theme(legend.position = "none",
      panel.grid.minor = element_blank(),
      title = element_text(family = "courier", face = "bold"))

p1_npg + p1_rickandmorty + p3_gsea
```

:::{.pt5}
Code source: [https://nanx.me/ggsci/articles/ggsci.html](https://nanx.me/ggsci/articles/ggsci.html)
:::

## `ggnewscale`

```{r}
#| fig-width: 12
#| fig-height: 4

library(ggnewscale)
# Equivalent to melt(volcano)
topography <- expand.grid(x = 1:nrow(volcano),
                          y = 1:ncol(volcano))
topography$z <- c(volcano)

# point measurements of something at a few locations
set.seed(42)
measurements <- data.frame(x = runif(30, 1, 80),
                           y = runif(30, 1, 60),
                           thing = rnorm(30))

ggplot(mapping = aes(x, y)) +
  geom_contour(data = topography, aes(z = z, color = stat(level))) +
  # Color scale for topography
  scale_color_viridis_c(option = "D") +
  # geoms below will use another color scale
  new_scale_color() +
  geom_point(data = measurements, size = 3, aes(color = thing)) +
  # Color scale applied to geoms added after new_scale_color()
  scale_color_viridis_c(option = "A")

```

:::{.pt5}
Code source: [https://github.com/eliocamp/ggnewscale/](https://github.com/eliocamp/ggnewscale/)
:::

## `ggbreak`

```{r}
#| fig-width: 12
#| fig-height: 4
library(ggbreak)

d <- data.frame(x = 1:20,
   y = c(rnorm(5) + 4, rnorm(5) + 20, rnorm(5) + 5, rnorm(5) + 22)
)
 
p1 <- ggplot(d, aes(y, x)) + geom_col(orientation="y") +
  scale_x_break(c(7, 17), scales = 1.5) + scale_x_break(c(18, 21), scales=2) +
  labs(title = "scale_x_break()") +
  theme_bw() +
  theme(legend.position = "none",
      panel.grid.minor = element_blank(),
      title = element_text(family = "courier", face = "bold")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

p2 <- ggplot(economics, aes(x=date, y = unemploy, colour = uempmed)) +
  geom_line() +
  scale_wrap(n=4) +
  labs(title = "scale_wrap()") +
  theme_bw() +
  theme(legend.position = "none",
      panel.grid.minor = element_blank(),
      title = element_text(family = "courier", face = "bold"))

d <- data.frame(
     x = 1:20,
     y = c(rnorm(5) + 4, rnorm(5) + 20, rnorm(5) + 5, rnorm(5) + 22)
 )

p3 <- ggplot(d, aes(x, y)) +
  geom_col() +
  scale_y_cut(breaks=c(7, 18), which=c(1, 3), scales=c(3, 0.5)) +  scale_y_cut(breaks=c(7, 18), which=c(1, 3), scales=c(3, 0.5), space=.5) +
  labs(title = "scale_y_cut()") +
  theme_bw() +
  theme(legend.position = "none",
      panel.grid.minor = element_blank(),
      title = element_text(family = "courier", face = "bold"))

p1 + p2 + p3
```
:::{.pt5}
Code source: [https://cran.r-project.org/web/packages/ggbreak/vignettes/ggbreak.html](https://cran.r-project.org/web/packages/ggbreak/vignettes/ggbreak.html)
:::


## New Stats

`ggstats` -- 

`ggdist` -- 

`ggpmisc` -- 


## New Facets

`geofacet` -- 

`ggh4x` -- 

`ggside` -- implemented as `geoms` but really adds facets


## New Themes

`ggthemes` -- 

`hrbrthemes` -- 

`ggokabeito` --


# Classifying by construct:
ALTERNATIVE INITALIZE PLOT FUNCTION

## Transforms Data

`waterfalls` --

## Returns ggplot2 Object

## Does Not Return ggplot2 Object

`survminer` -- 

`gge`

## New Components

`gggenomes` -- 

`ggformula`

## `gggenomes`

```{r}
library(gggenomes)
# to inspect the example data shipped with gggenomes
data(package="gggenomes")

gfull <- gggenomes(
  genes = emale_genes, seqs = emale_seqs, links = emale_ava,
  feats = list(emale_tirs, ngaros=emale_ngaros, gc=emale_gc)) |> 
  add_sublinks(emale_prot_ava) |>
  sync() + # synchronize genome directions based on links
  geom_feat(position="identity", size=6) +
  geom_seq() +
  geom_link(data=links(2)) +
  geom_bin_label() +
  geom_gene(aes(fill=name)) +
  geom_gene_tag(aes(label=name), nudge_y=0.1, check_overlap = TRUE) +
  geom_feat(data=feats(ngaros), alpha=.3, size=10, position="identity") +
  geom_feat_note(aes(label="Ngaro-transposon"), data=feats(ngaros),
      nudge_y=.1, vjust=0) +
  geom_wiggle(aes(z=score, linetype="GC-content"), feats(gc),
      fill="lavenderblush4", position=position_nudge(y=-.2), height = .2) +
  scale_fill_brewer("Genes", palette="Dark2", na.value="cornsilk3")

gfull
```

:::{.pt5}
Code source: [https://thackl.github.io/gggenomes/](https://thackl.github.io/gggenomes/)
:::

## Modular components

```{r}
#| layout-ncol: 2
g <- gggenomes(
  genes = emale_genes, seqs = emale_seqs, links = emale_ava,
  feats = list(emale_tirs, ngaros=emale_ngaros, gc=emale_gc)) |> 
  add_sublinks(emale_prot_ava) |>
  sync() + # synchronize genome directions based on links
  geom_seq(color = NA) +
  geom_bin_label(color = NA) +
  scale_fill_brewer("Genes", palette="Dark2", na.value="cornsilk3") + 
  theme(legend.position = "none",
        title = element_text(size = 8, family = "courier", face = "bold"))
g1 <- g + geom_feat(position="identity", size=6) +
  labs(title = "geom_feat()")
g2 <- g + geom_seq() + labs(title = "geom_seq()")
g3 <- g + geom_link(data=links(2)) + labs(title = "geom_link()")
g4 <- g + geom_bin_label(size = 2) + labs(title = "geom_bin_label()")
g5 <- g + geom_gene(aes(fill=name)) + labs(title = "geom_gene()")
g6 <- g + geom_gene_tag(aes(label=name), nudge_y=0.1, check_overlap = TRUE) + labs(title = "geom_gene_tag()")
g7 <- g + geom_feat(data=feats(ngaros), alpha=.3, size=10, position="identity") + labs(title = "geom_feat()")
g8 <- g + geom_feat_note(aes(label="Ngaro-transposon"), data=feats(ngaros), nudge_y=.1, vjust=0) +
   labs(title = "geom_feat_note()")
g9 <- g + geom_wiggle(aes(z=score, linetype="GC-content"), feats(gc), fill="lavenderblush4", position=position_nudge(y=-.2), height = .2) + labs(title = "geom_wiggle()")
gfull
g1 + g2 + g3 + g4 + g5 + g6 + g7 + g8 + g9
```

## Modular components

```{r}
g1 + g2 + g3 + g4 + g5 + g6 + g7 + g8 + g9
```


# Classifying by purpose:

## Specialized Plots

## Labels/Text

## Plot Positioning

## Interactivity

`ggiraph`


## Animation

`ggreveal` -- add parts incrementally

## Biostatistics

## Statistics Integration



## Altered Geoms

`ggpattern`


## Chart type built modularly

`ggalluvial` -- alluvial diagrams



## Non-cartesian extensions

(Or are these new coordinate systems?)

`ggraph` -- networks 

`ggtern` -- ternary diagrams (3 axes)


## Add stuff to plots

`ggpp` -- tables, inset plots, images

## Complete plots, nonmodular

`see` -- effect plots and others

`ggstats` -- `ggcoef_model()`, `ggcoef_table()`, `ggcoef_compare()`, `ggsurvey()`, `gglikert()`, `ggcascade()`